---
title: 'DATA605: Fundamentals of Computational Mathematics'
author: "Donald Butler"
date: "05/07/2022"
output:
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document: default
subtitle: Assignment 11
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regression Analysis

Using the cars dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook Chapter 3 (visualization, quality evaluation of the model, and residual analysis).

### Load cars dataset

The cars dataset was collected in the 1920s and contains the speed in miles per hour and the stopping distance in feet.

```{r}
summary(cars)
```

## Construct Linear Model

```{r}
cars.lm = lm(formula = dist ~ speed, data = cars)
plot(formula = dist ~ speed, data = cars, main = 'Speed vs Stopping Distance', xlab = 'Speed (mph)', ylab = 'Stopping Distance (ft)')
abline(cars.lm)

```

### Model Evaluation

```{r}
summary(cars.lm)
```

The p-value of 1.5e-12 is less than .05 which indicates that speed is a significant variable. The $R^2$ value indicates that variations in speed account for 65% of the variation in stopping distance. The inter-quartile range is consistent with the residual standard error which supports that the residuals are normally distributed.

### Residual Analysis

Viewing a histogram of the residuals shows that they are nearly normal centered at 0.

```{r}
hist(resid(cars.lm))
```

The residuals appear to have constant variability.

```{r}
plot(fitted(cars.lm),resid(cars.lm))
```

The Q-Q plot of residuals shows that they are distributed nearly normal. There are a few outliers at higher speeds, but not too concerning.

```{r}
qqnorm(resid(cars.lm))
qqline(resid(cars.lm))
```

