---
title: 'DATA605: Fundamentals of Computational Mathematics'
subtitle: 'Assignment 8'
author: 'Donald Butler'
date: '03/20/2022'

output:
    pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1.

A company buys 100 light bulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?

```{r}
bulbs <- 100
expected_lifetime <- 1000
rate <- 1/expected_lifetime
trials <- 50000

results <- array(dim = trials)
for (trial in 1:trials) {
  results[trial] <- min(rexp(bulbs, rate))
}

mean(results)
```

# Problem 2.

Assume that $X_1$ and $X_2$ are independent random variables, each having an exponential density with parameter $\lambda$. Show that $Z = X_1 - X_2$ has density $$f_Z(z) = (\frac{1}{2})\lambda e^{-\lambda |z|}$$.

$$f_{X_1}(x) = f_{X_2}(x) = \begin{cases}\lambda e^{-\lambda x}, &\text{if }x \ge 0, \\ 0, &\text{otherwise};\end{cases}$$

$$f_{X_1}(x_1)*f_{X_2}(x_2) = \lambda^2e^{-\lambda(x_1+x_2)}$$

Since $Z = X_1 - X_2$, $x_1 = z+x_2$. $\lambda^2e^{-\lambda((z+x_2)+x_2)}=\lambda^2e^{-\lambda(z+2x_2)}$.

If $z \ge 0$, $x_2 \ge 0$, so $\int_0^{\infty} \lambda^2e^{-\lambda(z+2x_2)} = \frac{\lambda}{2}e^{-\lambda z}$.

Since $X_1$ and $X_2$ are independent, the distribution of $X_1 - X_2$ is the same as $X_2 - X_1$. This means that $f_Z(z) = f_Z(-z)$, so $$f_Z(z) = \frac{\lambda}{2}e^{-\lambda |z|}$$


# Problem 3.

Let $X$ be a continuous random variable with mean $\mu = 10$ and variance $\sigma^2 = 100/3$. Using Chebyshev's Inequality, find an upper bound for the following probabilities.

> **Theorem 8.3 (Chebyshev Inequality)**
  Let $X$ be a continuous random variable with density function $f(x)$. Suppose $X$ has a finite expected value $\mu = E(X)$ and finite variance $\sigma^2 = V(X)$. Then for any positive number $\epsilon > 0$ we have $$P(|X-\mu|\ge \epsilon)\le \frac{\sigma^2}{\epsilon^2}$$.


## a.

$P(|X-10| \ge 2)$.

```{r}
V = 100/3
e = 2
V/e^2
```

## b.

$P(|X-10| \ge 5)$.

```{r}
V = 100/3
e = 5
V/e^2
```

## c.

$P(|X-10| \ge 9)$.

```{r}
V = 100/3
e = 9
V/e^2
```

## d.

$P(|X-10| \ge 20)$.

```{r}
V = 100/3
e = 20
V/e^2
```

