---
title: 'DATA605: Fundamentals of Computational Mathematics'
subtitle: 'Assignment 3'
author: 'Donald Butler'
date: '02/12/2022'

output:
    pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Set 1

1. What is the rank of the matrix A?

$$A = \begin{bmatrix}1 & 2 & 3 & 4\\-1 & 0 & 1 & 3\\0 & 1 & -2 & 1\\5 & 4 & -2 & -3\end{bmatrix}$$

The rank of a matrix is the maximum number of linearly independent columns of $A$, which can be found by counting the number of pivot columns in the reduced row echelon form of the matrix.

```{r}
library(pracma)
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), 4, 4, byrow = TRUE)
rref(A)
```
The $rref(A)$ has 4 pivot columns, so the rank of A is 4.

2. Given an $mxn$ matrix where $m > n$, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

The maximum rank for a matrix is the smaller of the number of columns or the number of rows, since $m > n$, the maximum rank is $n$. If the matrix is non-zero, the rank must be at least 1, which is achieved if the columns are all multiples of each other.

3. What is the rank of matrix B?

$$B = \begin{bmatrix}1 & 2 & 1\\3 & 6 & 3 \\2 & 4 & 2\end{bmatrix}$$

Just looking at $B$, you can tell that the rank is 1 because the columns are multiples of each other, $C_2 = 2C_1$ and $C_3 = C_1$.

# Problem Set 2

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

$$A = \begin{bmatrix}1 & 2 & 3 \\0 & 4 & 5 \\0 & 0 & 6\end{bmatrix}$$

```{r}
A = matrix(c(1,2,3,0,4,5,0,0,6),3,3,byrow = TRUE)
```

\begin{align}Ax &= \lambda{}x \\ 0 &= \lambda{}x-Ax \\ 0 &= \lambda{}I_{n}x-Ax \\ 0 &= (\lambda{}I_{n}-A)x\end{align}

This has solutions when $det(\lambda{}I-A) = 0$.

\begin{align}\begin{vmatrix}\lambda{}-1 & -2 & -3 \\ 0 & \lambda{}-4 & -5 \\ 0 & 0 & \lambda{}-6\end{vmatrix}&=(\lambda{}-1)\begin{vmatrix}\lambda{}-4 & -5 \\ 0 & \lambda{}-6\end{vmatrix}-0\begin{vmatrix}-2 & -3 \\ 0 & \lambda{}-6\end{vmatrix}+0\begin{vmatrix}-2 & -3 \\ \lambda{}-4 & -5\end{vmatrix} \\ &=(\lambda{}-1)((\lambda{}-4)(\lambda{}-6)-(-5*0)) \\ &=(\lambda{}-1)(\lambda{}-4)(\lambda{}-6) \end{align}

The characteristic polynomial is $\rho_A(x) = (x-1)(x-4)(x-6)$.

For $\lambda{}=1$

```{r}
lambda = 1
rref(lambda*diag(3)-A)
```

$\begin{bmatrix}0 & 1 & 0 \\ 0 & 0 & 1 \\ 0 & 0 & 0\end{bmatrix}\begin{bmatrix}v_1 \\ v_2 \\ v_3\end{bmatrix}=0$

$v_2=0 \\ v_3 = 0$

$\xi_A(1) = \left \langle \begin{bmatrix}1 \\ 0 \\ 0\end{bmatrix} \right \rangle$

For $\lambda{}=4$

```{r}
lambda = 4
rref(lambda*diag(3)-A)
```

$\begin{bmatrix}1 & -\frac{2}{3} & 0 \\ 0 & 0 & 1 \\ 0 & 0 & 0\end{bmatrix}\begin{bmatrix}v_1 \\ v_2 \\ v_3\end{bmatrix}=0$

$v_1-\frac{2}{3}v_2=0; v_1 = \frac{2}{3}v_2 \\ v_3 = 0$

$\xi_A(4) = \left \langle \begin{bmatrix}2 \\ 3 \\ 0\end{bmatrix} \right \rangle$

For $\lambda{}=6$  

```{r}
lambda = 6
rref(lambda*diag(3)-A)
```

$\begin{bmatrix}1 & 0 & -\frac{8}{5} \\ 0 & 1 & -\frac{5}{2} \\ 0 & 0 & 0\end{bmatrix}\begin{bmatrix}v_1 \\ v_2 \\ v_3\end{bmatrix}=0$

$v_1-\frac{8}{5}v_3=0; v_1 = \frac{8}{5}v_3 \\ v_2-\frac{5}{2}v_3=0; v_2 = \frac{5}{2}v_3$

$\xi_A(6) = \left \langle \begin{bmatrix}16 \\ 25 \\ 10\end{bmatrix} \right \rangle$

