---
title: 'DATA605: Fundamentals of Computational Mathematics'
author: "Donald Butler"
date: "05/08/2022"
output:
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document: default
subtitle: Assignment 15
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Find the equation of the regression line for the given points.

$$(5.6,8.8),(6.3,12.4),(7,14.8),(7.7,18.2),(8.4,20.8)$$


```{r}
df <- data.frame(x = c(5.6,6.3,7,7.7,8.4), y = c(8.8,12.4,14.8,18.2,20.8))
df.lm = lm(formula = y ~ x, data = df)
plot(df)
abline(df.lm)
```

```{r}
summary(df.lm)
```

The equation for the regression line is: $y = 4.257x - 14.8$.

## Problem 2

Find all local maxima, local minima, and saddle points for the function given below.

$$f(x,y) = 24x - 6xy^2 - 8y^3$$

Use first partial derivatives to identify critical points.

$$f_x = 24 - 6y^2$$
Evaluate $f_x = 0$.

$\begin{aligned}
0 & = 24 - 6y^2 \\
6y^2 & = 24 \\
y^2 & = 4 \\
y & = \pm 2
\end{aligned}$

$$f_y = -12xy - 24y^2$$

Evaluate $f_y = 0$.

$\begin{aligned}
0 & = -12xy - 24y^2 \\
12xy & = -24y^2 \\
x & = -2y
\end{aligned}$

This gives the critical points, $(4, -2, 64)$ and $(-4, 2,-64)$.

Use the second derivative test to determine the type of critical points.

$$\begin{aligned}
f_{xx} & = 0 \\
f_{xy} & = -12y \\
f_{yy} & = -12x - 48y
\end{aligned}$$

Evaluate $D = f_{xx}(4,-2)f_{yy}(4,-2)-f_{xy}^2(4,-2)$ and $D = f_{xx}(-4,2)f_{yy}(-4,2)-f_{xy}^2(-4,2)$.

$\begin{aligned}
D & = f_{xx}(4,-2)f_{yy}(4,-2)-f_{xy}^2(4,-2) \\
& = 0 * (-48 + 96) - (24^2) \\
& = -(24^2) \\
& < 0
\end{aligned}$

The first critical point $(4, -2, 64)$ is a saddle point.

Evaluate $D = f_{xx}(-4,2)f_{yy}(-4,2)-f_{xy}^2(-4,2)$.

$\begin{aligned}
D & = f_{xx}(-4,2)f_{yy}(-4,2)-f_{xy}^2(-4,2) \\
& = 0 * (48 - 96) - ((-24)^2) \\
& = -(24^2) \\
& < 0
\end{aligned}$

The second critical point $(-4, 2, -64)$ is also saddle point.

## Problem 3

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for $x$ dollars and the "name" brand for $y$ dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

### Part 1

Find the revenue function $R(x,y)$.

$\begin{aligned}
R(x,y) & = x (81 - 21x + 17y) + y (40 + 11x - 23y) \\
& = 81x - 21x^2 + 17xy + 40y + 11xy - 23y^2 \\
& = 81x - 21x^2 + 28xy + 40y - 23y^2
\end{aligned}$

### Part 2

What is the revenue if she sells the "house" brand for \$2.30 and the "name" brand for \$4.10?

```{r}
R <- function(x,y){81*x - 21*x^2 + 28*x*y + 40*y - 23*y^2}
R(2.30,4.10)
```

## Problem 4

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x,y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$, where $x$ is the number of units produced in Los Angeles and $y$ is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

Since we need to produce 96 units, $x + y = 96$. Substitute $y = 96 - x$ in $C(x,y)$ and evaluate in terms of just $x$.

$\begin{aligned}
C(x) & = \frac{1}{6}x^2 + \frac{1}{6}(96-x)^2 + 7x + 25(96-x) + 700 \\
& = \frac{1}{6}x^2 + \frac{1}{6}(9216-192x+x^2) + 7x + 2400 - 25x + 700 \\
& = \frac{1}{3}x^2 - 50x + 4636 
\end{aligned}$

Evaluate the first derivative to find the critical points.

$\begin{aligned}
C'(x) & = \frac{2}{3}x - 50 \\
0 & = \frac{2}{3}x - 50 \\
50 & = \frac{2}{3}x \\
75 & = x
\end{aligned}$

Since $x = 75$, $y = 96 - 75 = 21$. Produce 75 units in Los Angeles and 21 units in Denver.

## Problem 5

Evaluate the double integral on the given region.

$$\iint_R (e^{8x+3y})dA; R:2\le x \le 4 \text{ and } 2 \le y \le 4$$

$\begin{aligned}
\int_2^4 \int_2^4 e^{8x+3y}dx dy & = \int_2^4 \int_2^4 e^{8x}e^{3y}dx dy \\
& = \int_2^4 e^{8x}dx * \int_2^4 e^{3y}dy \\
& = \frac{e^{8x}}{8} \Big|_2^4 * \frac{e^{3y}}{3} \Big|_2^4 \\
& = \frac{e^{32}-e^{16}}{8} * \frac{e^{12}-e^{6}}{3} \\
& = \frac{1}{24} (e^{44}-e^{38}-e^{28}+e^{22})
\end{aligned}$
