---
title: 'DATA605: Fundamentals of Computational Mathematics'
subtitle: 'Assignment 7'
author: 'Donald Butler'
date: '03/12/2022'

output:
    pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1.

Let $X_1$, $X_2$, ..., $X_n$ be $n$ mutually independent random variables, each of which is uniformly distributed on the integers from $1$ to $k$. Let $Y$ denote the minimum of the $X_i$s. Find the distribution of $Y$.

The distribution of $Y$ is discrete on the interval $[1,k]$ and can be calculated by finding the probability that the minimum value of the $X_i$s each value. There are $k^n$ ways of selecting the $X_i$ values. For $1 \le j \le k$, the number of ways that each $X_i$ can be selected so that the minimum value is $j$, is the number of ways that $X_i \ge j$ minus the ways that $X_i > j$. So, the distribution of $Y$ is:
$$\frac{(k-j+1)^n - (k-j)^n}{k^n}, 1 \le j \le k$$

# Problem 2.

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years.

## Part a. (geometric)

What is the probability that the machine will fail after 8 years?

```{r}
p <- .1

(prob_fail_after_8 <- (1-p)^8)
(ev <- 1/p)
(sd <- sqrt((1-p)/p^2))
```

## Part b. (exponential)

What is the probability that the machine will fail after 8 years?

```{r}
p <- .1

(prob_fail_after_8 <- 1 - pexp(8,p))
(ev <- 1/p)
(sd <- sqrt(1/p^2))
```

## Part c. (binomial)

What is the probability that the machine will fail after 8 years?

```{r}
p <- .1

(prob_fail_after_8 <- dbinom(0,8,p))
(ev <- 8 * p)
(sd <- sqrt(8 * p * (1-p)))
```

## Part d. (poisson)

What is the probability that the machine will fail after 8 years?

```{r}
p <- .1

(prob_fail_after_8 <- dpois(0,8 * p))
(ev <- 8 * p)
(sd <- sqrt(8 * p))
```
