---
title: 'DATA605: Fundamentals of Computational Mathematics'
author: "Donald Butler"
date: "03/27/2022"
output:
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document: default
subtitle: Assignment 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1.

The price of one share of stock in the Pilsdorff Beer Company is given by $Y_n$ on the nth day of the year. Finn observes that the differences $X_n = Y_{n+1}-Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma^2 = 1/4$. If $Y_1 = 100$, estimate the probability the $Y_365$ is:

The price of a stock on day N is the sum of the previous days differences. 


## 1a. $\ge 100$.


```{r}
n <- 364
mu <- 0
var <- 1/4

Sn <- 100 - 100 #Y_365 - Y_1
sd <- sqrt(n * var)

pnorm(Sn, mu, sd, lower.tail = FALSE)

```


## 1b. $\ge 110$.

```{r}
n <- 364
mu <- 0
var <- 1/4

Sn <- 110 - 100 #Y_365 - Y_1
sd <- sqrt(n * var)

pnorm(Sn, mu, sd, lower.tail = FALSE)

```


## 1c. $\ge 120$.

```{r}
n <- 364
mu <- 0
var <- 1/4

Sn <- 120 - 100 #Y_365 - Y_1
sd <- sqrt(n * var)

pnorm(Sn, mu, sd, lower.tail = FALSE)

```

# Problem 2.

Calculate the expected value and variance of the binomial distribution using the moment generating function.

The moment generating function for the binomial distribution is $g(t) = (pe^t + (1-p))^n$. To calculate the mean and variance, we need the first and second derivatives of $g(t)$. 
$$g'(t) = n(pe^t + (1-p))^{n-1}(pe^t)$$
$$g''(t) = n(n-1)(pe^t + (1-p))^{n-2}(pe^t)^2+n(pe^t + (1-p))^{n-1}(pe^t)$$

$$\begin{aligned}
\mu & = g'(0) \\
& = n(pe^0+(1-p))^{n-1}(pe^0) \\
& = n(p+(1-p))^{n-1}(p) \\
& = n(1^{n-1})p \\
& = np
\end{aligned}$$

$$\begin{aligned}
\sigma^2 & = g''(0) - \mu^2 \\
& = n(n-1)(pe^0+(1-p))^{n-2}(pe^0)^2+n(pe^0 + (1-p))^{n-1}(pe^0) - (np)^2\\
& = (n^2-n)(p+(1-p))^{n-2}p^2 + n(p + (1-p))^{n-1}p - (np)^2 \\
& = (np)^2 - np^2 +np - (np)^2 \\
& = np-np^2 \\
& = np(1-p) \\
\end{aligned}$$


# Problem 3.

Calculate the expected value and variance of the exponential distribution using the moment generating function.

$$g(t) = \frac{1}{1-t/\lambda}$$
$$g'(t) = (1/\lambda)(1-t/\lambda)^{-2}$$

$$g''(t) = (2/\lambda)^2(1-t/\lambda)^{-3}$$

$$\begin{aligned}
\mu & = g'(0) \\
& = 1 / \lambda
\end{aligned}$$


$$\begin{aligned}
\sigma^2 & = g''(0) - \mu^2 \\
& = (2/\lambda)^2 - (1/\lambda)^2 \\
& = 1/\lambda^2
\end{aligned}$$





